---
title: "Lab 7 - Space Use and Resource Selection"
subtitle: "ECOL 620 - Applications in Landscape Ecology"
author: "George Woolsey"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document:
    toc: true
    toc_float: true
    # code_folding: hide
editor_options: 
  chunk_output_type: console
knit: (function(inputFile, encoding){ 
    out_dir <- '../';
    rmarkdown::render(inputFile, encoding = encoding, output_file=file.path(dirname(inputFile), out_dir, 'index.html'));
    file.copy(from = "..//index.html", to = '..///data//lab7_george_woolsey.html', overwrite = TRUE)
  })
---

# Setup

```{r, include=FALSE, warning=F, message=F}
# knit options
knitr::opts_chunk$set(
  echo = TRUE
  , warning = FALSE
  , message = FALSE
  # , results='hide'
  , fig.width = 10
  , fig.height = 7
)
```

```{r}
# bread-and-butter
library(tidyverse)
library(lubridate)
library(viridis)
library(scales)
library(latex2exp)
# visualization
library(kableExtra)
library(cowplot)
library(RColorBrewer)
library(ggnewscale) # add second scale color/fill
# spatial analysis
library(terra)
library(adehabitatLT)     #for trajectory information; version 0.3.23 used
library(adehabitatHR)     #Home range estimation; version 0.4.15 used
library(adehabitatHS)     #for selection ratios; version 0.3.13 used
library(maptools) ## one of these adehabitat needs this
# do i have to???
# library(raster)           #for raster covariate data; version 2.6-7 used
# library(rgdal)            #for reading different types of GIS files; version 1.3-4 used
# set seed
set.seed(11)
```

# Rules

1. Using the provided R Markdown template, answer all questions and show your R code where necessary. Note, some questions are just plain text written answers.

2. Complete your assignment using the R markdown file and submit individual assignments to Canvas. Knit your script and submit an .html file on Canvas. Please use  the following naming convention: lab5_firstname_lastname.html (ex. lab7_kyle_horton.html).  Note, we will not grade labs in any other format.

---

# Florida panther home range

```{r, include=FALSE, eval=FALSE}
...notes...

local convex hull map might be difficult...check for shapefile uploaded to canvas
  is the issue with the holes in the polygons? can use st_make_valid?

resource selection has various levels...focus on type 2 and type 3...trying to figure out where the individuals are, are they selecting preferentially basee on what's available?

type 2 is at broad landscape scale, type 3 narrows down to home range
```


## Load Data

```{r, results='hide', fig.show='hide'}
# landcover source: fwc/fnai
land <- terra::rast("../data/panther_landcover.tif")
reproject_crs <- terra::crs(land)
terra::freq(land)
# Add panther data
panthers <- sf::st_read("../data/panthers.shp") |> 
  sf::st_transform(reproject_crs) |> 
  dplyr::select(-c(X,Y)) |> 
  dplyr::mutate(CatID = as.factor(CatID))
#inspect
str(panthers)
# counts
panthers |> 
  sf::st_drop_geometry() |> 
  dplyr::count(CatID)
panthers |> 
  sf::st_drop_geometry() |> 
  dplyr::count(AgeClass)
panthers |> 
  sf::st_drop_geometry() |> 
  dplyr::count(CatID,AgeClass)
# plot
ggplot() + 
  geom_raster(
    data = land |> as.data.frame(xy=T) |> dplyr::rename(covertyp=3)
    , mapping = aes(x=x,y=y,fill=covertyp)
  ) +
  geom_sf(data = panthers, mapping = aes(color = CatID)) +
  scale_fill_viridis_c(alpha = 0.8) +
  scale_color_viridis_d(option = "magma") +
  theme_light()
```

Reclassify the land cover data based on classification matrix

```{r, results='hide', fig.show='hide'}
# reclass
# load reclassification to resistance surface
landcover_reclass <- read.table("../data/landcover reclass.txt", header = T) |>  
  dplyr::rename_with(tolower)
landcover_reclass |>  dplyr::glimpse()
# create landcover surface raster
rcl_matrix <- landcover_reclass |>  dplyr::select(landcover, changeto) |>  as.matrix()
land_sub <- terra::classify(land, rcl = rcl_matrix)
# investigate
land |>  terra::unique() |>  nrow()
land_sub |>  terra::unique() |>  nrow()
plot(land_sub)

#create forested wetlands layer
wetforest <- land_sub
terra::values(wetforest) <- 0
wetforest[land_sub %in% c(9,11)] <- 1
plot(wetforest)
#create forested uplands layer
dryforest <- land_sub
terra::values(dryforest) <- 0
dryforest[land_sub %in% c(10,12)] <- 1
plot(dryforest)
# create moving window calculation of different forest types
fw <- terra::focalMat(land_sub, d = 5000, type = "circle")
dry.focal <- terra::focal(dryforest, w = fw, fun = "sum", na.rm=T)
wet.focal <- terra::focal(wetforest, w = fw, fun = "sum", na.rm=T)
plot(dry.focal)
plot(wet.focal)
```

```{r, warning=FALSE, message=FALSE, echo=FALSE, include=FALSE}
remove(list = ls()[grep("_temp",ls())])
gc()
```

## Question 1

Within the "panthers" SpatialPointsDataFrame, there are three columns of information. Describe what each represents. (3 pts)

<span style="color: teal;">
The "panthers.shp" data includes radiotelemetry data from Florida panthers (*Puma concolor coryi*) in south Florida with the columns **`r paste0(names(panthers |> sf::st_drop_geometry()), collapse = ", ")`**. Data come from six panthers identified by the column `CatID`, three subadults and three adults denoted in the data column `AgeClass`, that were tracked during 2006 with dates identified by the column `Juldate`.
</span>

## Question 2

Describe what lines 102-104 are achieving. What is "fw" and how is it used in the "focal" function? (4 pts)

```{r, eval=FALSE, include=TRUE}
fw <- raster::focalWeight(land_sub, 5000, 'circle')
dry.focal <- raster::focal(dryforest, w=fw, fun="sum", na.rm=T)
wet.focal <- raster::focal(wetforest, w=fw, fun="sum", na.rm=T)
```

<span style="color: teal;">
The `raster::focalWeight` function calculates a "moving window" weight matrix in which the values add up to one, named `fw`, in a 5 km radius (5,000 m) for use in the `raster::focal` function. The `raster::focal` function calculates the focal ("moving window") values for the neighborhood of focal cells using a matrix of weights and sums the weighted amount of each forest type within the 5 km radius.
</span>

## Question 3

Plot the minimum convex polygon for panther #137. Show the raw locations (points) and the 50%, 75%, and 95% minimum convex polygons. See the example on Canvas for reference.  (5 pts)

```{r}
# take panthers and select the column of the data frame that 
  # includes the animal identification number (CatID) to 
  # calculate Minimum Convex Polygons with the mcp function
# :'( :'( :'(  :'( :'( :'(  :'( :'( :'(  :'( :'( :'( 
# have to convert from sf to sp object to use mcp
sp_panthers_temp <- sf::as_Spatial(panthers)
# function to pass to mcp
my_mcp_fn <- function(pct) {
  adehabitatHR::mcp(
    sp_panthers_temp[,"CatID"]
    , percent = pct
  ) |> 
  sf::st_as_sf()
}
# pcts for minimum convex polys
pct_list <- c(95,75,50)
# create sf object with all pcts in one for plotting
min_convex_polys <- pct_list |> 
  purrr::map(my_mcp_fn) |> 
  dplyr::bind_rows(.id = "column_label") |> 
  dplyr::mutate(
    pct_grp = factor(
      column_label
      , labels = paste0(pct_list,"%")
      , ordered = T
    )
  )
```

Plot for panther #137

```{r}
panther_filter_temp <- "137"
# create df of landcover reclass raster
land_sub_df_temp <- land_sub |> 
  terra::crop(
    panthers |> 
      dplyr::filter(CatID == panther_filter_temp) |> 
      sf::st_bbox() |> 
      sf::st_as_sfc() |> 
      sf::st_buffer(5000) |> 
      terra::vect()
  ) |> 
  as.data.frame(xy=T) |> 
  dplyr::rename(landcover_type=3) |> 
  dplyr::left_join(
    landcover_reclass |> 
      dplyr::count(changeto,description2) 
    , by = c("landcover_type"="changeto")
  ) |> 
  dplyr::mutate(
    landcover_class = as.factor(description2)
  )
#plot
ggplot() +
  geom_raster(
    data = land_sub_df_temp
    , mapping = aes(x=x,y=y,fill=landcover_class)
  ) +
  geom_sf(
    data = min_convex_polys |> dplyr::filter(id == panther_filter_temp)
    , mapping = aes(color = pct_grp)
    , fill = "gray"
    , alpha = 0.1
    , lwd = 2
  ) +
  geom_sf(
    data = panthers |> dplyr::filter(CatID == panther_filter_temp)
    , color = "black"
    , fill = "gray85"
    , shape = 21
  ) +
  scale_fill_viridis_d(option = "turbo", alpha = 0.8) +
  scale_color_manual(
    values = RColorBrewer::brewer.pal(n = length(pct_list)*2, name = "Greys") |> tail(length(pct_list))
  ) +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0)) +
  labs(
    subtitle = paste0("Radiotelemetry locations and minimum convex polygons for panther #", panther_filter_temp)
    , x = "Longitude"
    , y = "Latitude"
    , fill = "Landcover Class"
    , color = "Minimum convex polygons\nHome Range %"
  ) +
  theme_light() +
  theme(
    legend.position = c(0.12, 0.2)
    , legend.key.size = unit(0.25, 'cm')
    , legend.text = element_text(size = 7)
    , legend.title = element_text(size = 7)
    , axis.title = element_text(size = 7)
    , axis.text.y = element_text(size=6)
    , axis.text.x = element_text(size=6)
    , panel.border = element_blank()
    , plot.subtitle = element_text(size=9)
    , panel.grid = element_blank()
  )

```

```{r, warning=FALSE, message=FALSE, echo=FALSE, include=FALSE}
remove(list = ls()[grep("_temp",ls())])
gc()
```

## Question 4

Plot the minimum convex polygon for all panthers overlaid on the reclassified landcover raster. Show the 50% and 95% minimum convex polygons. See the example on Canvas for reference. (6 pts)

```{r}
# create df of landcover reclass raster
land_sub_df <- land_sub |> 
  as.data.frame(xy=T) |> 
  dplyr::rename(landcover_type=3) |> 
  dplyr::left_join(
    landcover_reclass |> 
      dplyr::count(changeto,description2) 
    , by = c("landcover_type"="changeto")
  ) |> 
  dplyr::mutate(
    landcover_class = as.factor(description2)
  )
# filter for only some pcts
pct_list_temp <- c("50%","95%")
set.seed(11)
#plot
ggplot() +
  geom_raster(
    data = land_sub_df
    , mapping = aes(x=x,y=y,fill=landcover_class)
  ) +
  scale_fill_viridis_d(option = "turbo", alpha = 0.8) +
  labs(fill = "Landcover Class") +
  # start a new scale fill for points
  ggnewscale::new_scale_fill() +
  geom_sf(
    data = panthers
    , mapping = aes(fill = CatID)
    , color = "black"
    , shape = 21
    , show.legend = F
    , size = 1
  ) +
  scale_fill_viridis_d(option = "magma") +
  # plot min_convex_polys
  geom_sf(
    data = min_convex_polys |> dplyr::filter(pct_grp %in% pct_list_temp)
    , mapping = aes(color = pct_grp)
    , fill = "gray"
    , alpha = 0.2
    , lwd = 1
  ) +
  scale_color_manual(
    values = RColorBrewer::brewer.pal(n = length(pct_list_temp)*2, name = "Greys") |> tail(length(pct_list_temp))
  ) +
  # labels
  ggrepel::geom_text_repel(
    data = 
      min_convex_polys |> 
        dplyr::filter(pct_grp == "95%") |> 
        sf::st_centroid() |> 
        dplyr::mutate(id=paste0("#",id))
    , aes(label=id, geometry = geometry)
    , stat = "sf_coordinates"
    , size = 3.5
    , color = "black"
    , fontface = "bold.italic"
    , seed = 11
    , nudge_x = 100
  ) +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0)) +
  labs(
    subtitle = paste0("Radiotelemetry locations and minimum convex polygons for panthers")
    , x = "Longitude"
    , y = "Latitude"
    , color = "Minimum convex polygons\nHome Range %"
  ) +
  theme_light() +
  theme(
    legend.position = c(0.12, 0.2)
    , legend.key.size = unit(0.25, 'cm')
    , legend.text = element_text(size = 7)
    , legend.title = element_text(size = 7)
    , axis.title = element_text(size = 7)
    , axis.text.y = element_text(size=6)
    , axis.text.x = element_text(size=6)
    , panel.border = element_blank()
    , plot.subtitle = element_text(size=9)
    , panel.grid = element_blank()
  )

```

## Question 5

For all panthers, create a four-panel plot showing the home ranges as estimated by: minimum convex polygon, fixed bivariate kernel home range, local convex hull home range, and Brownian bridge. For each approach, plot the 95% contours. Hint, for the local convex hull home range, I used the "LoCoH.k" function with a k of 14. (12 pts). 

### minimum convex polygon

```{r, fig.show='hide'}
#plot
plt_minimum_convex_polygon <- ggplot() +
  geom_raster(
    data = land_sub_df
    , mapping = aes(x=x,y=y,fill=landcover_class)
  ) +
  geom_sf(data = panthers, mapping = aes(color = CatID), alpha = 0.5, show.legend = F) +
  # plot home range
  geom_sf(
    data = min_convex_polys |> dplyr::filter(pct_grp == "95%")
    , mapping = aes(color = id)
    , fill = "gray"
    , alpha = 0.1
    , lwd = 1
    , show.legend = F
  ) +
  scale_color_viridis_d(option = "magma") +
  scale_fill_viridis_d(option = "turbo", alpha = 0.8) +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0)) +
  labs(
    subtitle = paste0("minimum convex polygon home range")
    , x = "Longitude"
    , y = "Latitude"
    , color = "Cat ID"
    , fill = "Landcover Class"
  ) +
  theme_light() +
  theme(
    plot.margin = margin(t=4,1,1,1, "lines") 
    , legend.direction = "horizontal"
    , legend.position = c(0.5, 1.15)
    , legend.key.size = unit(0.25, 'cm')
    , legend.text = element_text(size = 6)
    , legend.title = element_text(size = 6)
    , axis.title = element_text(size = 7)
    , axis.text.y = element_text(size=6)
    , axis.text.x = element_text(size=6)
    , panel.border = element_blank()
    , plot.subtitle = element_text(size=9)
    , panel.grid = element_blank()
  ) 
plt_minimum_convex_polygon

```

### fixed bivariate kernel home range

this code was updated to use `sp` and `raster` only when a function required for input...otherwise using `sf` and `terra`

```{r, fig.show='hide'}
# :'( :'( :'(  :'( :'( :'(  :'( :'( :'(  :'( :'( :'( 
# have to convert from sf to sp object to use mcp
sp_panthers_temp <- sf::as_Spatial(panthers)
# fixed bivariate kernel home range
fixed_bivariate_kernel <- adehabitatHR::kernelUD(
    xy = sp_panthers_temp[,"CatID"]
    , grid = 500
    , h = "href"
    , kern = "bivnorm"
    , same4all = T
  ) |> 
  estUDm2spixdf() |> 
  sf::st_as_sf()
# transform wide to long (id by cat)
fixed_bivariate_kernel_sf <- 1:length(fixed_bivariate_kernel |> sf::st_drop_geometry()) |> 
  purrr::map(function(x){
    ct <- names(fixed_bivariate_kernel)[x]
    return(
      fixed_bivariate_kernel[x] |> 
        dplyr::rename(density=1) |> 
        dplyr::mutate(CatID = substr(ct, 2, nchar(ct)))
    )
  }) |> 
  dplyr::bind_rows() |> 
  dplyr::group_by(CatID) |>
  dplyr::mutate(
    density_scale = scale(density)
  ) |> 
  dplyr::ungroup()

# function
plt_fixed_bivariate_kernel_fn <- function(catid) {
  # one cat raster
  rast_temp <- dplyr::bind_cols(
      fixed_bivariate_kernel_sf |> dplyr::filter(CatID==catid) |> sf::st_coordinates(geometry)
      , fixed_bivariate_kernel_sf |> 
        sf::st_drop_geometry() |> 
        dplyr::filter(CatID==catid) |> 
        dplyr::select(density_scale)
    ) |> 
    terra::rast(
      type = "xyz"
      , crs = (fixed_bivariate_kernel_sf |> terra::vect() |> terra::crs()) 
    ) |>
    terra::crop(
      panthers |> 
        dplyr::filter(CatID == catid) |> 
        sf::st_bbox() |> 
        sf::st_as_sfc() |> 
        sf::st_buffer(6000) |> 
        terra::vect()
    ) |> 
    as.data.frame(xy=T) |> 
    rename(density=3)
  # qtile
  qtile_temp <- quantile(rast_temp$density, 0.95) |> as.numeric()
  return(
  ggplot() + 
    geom_sf(
      data = panthers |> dplyr::filter(CatID == catid)
      , color = "black"
      , fill = "gray85"
      , shape = 21
    ) +
    geom_raster(data = rast_temp, mapping=aes(x=x,y=y,fill=density)) + 
    geom_contour(data = rast_temp, mapping=aes(x=x,y=y,z=density), breaks = qtile_temp, color = "black") + 
    scale_fill_viridis_c(
      option="inferno"
      , direction=-1
      , alpha = 0.8
      , limits = c(0, max(rast_temp$density))
      , oob = scales::censor
      , na.value = "transparent"
    ) +
    scale_x_continuous(expand = c(0, 0)) +
    scale_y_continuous(expand = c(0, 0)) +
    labs(
      subtitle = paste0("fixed bivariate kernel home range for panther #", catid)
      , x = "Longitude"
      , y = "Latitude"
      , color = "density"
    ) +
    theme_light() +
    theme(
      legend.position = c(0.12, 0.2)
      , legend.key.size = unit(0.25, 'cm')
      , legend.text = element_text(size = 6)
      , legend.title = element_text(size = 6)
      , axis.title = element_text(size = 7)
      , axis.text.y = element_text(size=6)
      , axis.text.x = element_text(size=6)
      , panel.border = element_blank()
      , plot.subtitle = element_text(size=8)
      , panel.grid = element_blank()
    )
  )
}

plt_fixed_bivariate_kernel_fn("100")
```

#### Using old methods

```{r, fig.show='hide'}
# :'( :'( :'(  :'( :'( :'(  :'( :'( :'(  :'( :'( :'( 
sp_panthers_temp <- sf::as_Spatial(panthers)
fixed_bivariate_kernel_old <- adehabitatHR::kernelUD(
    xy = sp_panthers_temp[,"CatID"]
    , h = "href"
    , kern = "bivnorm"
  )
kernel.95 <- adehabitatHR::getverticeshr(fixed_bivariate_kernel_old, percent=95) |> 
  sf::st_as_sf() |> 
  dplyr::rename(CatID = id)
# plot
#plot
plt_fixed_bivariate_kernel <- ggplot() +
  geom_raster(
    data = land_sub_df
    , mapping = aes(x=x,y=y,fill=landcover_class)
    , show.legend = F
  ) +
  geom_sf(data = panthers, mapping = aes(color = CatID), alpha = 0.5, show.legend = F) +
  # plot home range
  geom_sf(
    data = kernel.95
    , mapping = aes(color = CatID)
    , fill = "gray"
    , alpha = 0.1
    , lwd = 1
  ) +
  scale_color_viridis_d(option = "magma") +
  scale_fill_viridis_d(option = "turbo", alpha = 0.8) +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0)) +
  labs(
    subtitle = paste0("fixed bivariate kernel home range")
    , x = "Longitude"
    , y = "Latitude"
    , color = "Cat ID"
    , fill = "Landcover Class"
  ) +
  theme_light() +
  theme(
    plot.margin = margin(t=4,1,1,1, "lines") 
    , legend.direction = "horizontal"
    , legend.position = c(0.5, 1.15)
    , legend.key.size = unit(0.25, 'cm')
    , legend.text = element_text(size = 6)
    , legend.title = element_text(size = 6)
    , axis.title = element_text(size = 7)
    , axis.text.y = element_text(size=6)
    , axis.text.x = element_text(size=6)
    , panel.border = element_blank()
    , plot.subtitle = element_text(size=9)
    , panel.grid = element_blank()
  ) 
plt_fixed_bivariate_kernel
```

### local convex hull home range

Hint, for the local convex hull home range, I used the "LoCoH.k" function with a k of 14.

```{r, fig.show='hide'}
# :'( :'( :'(  :'( :'( :'(  :'( :'( :'(  :'( :'( :'( 
# local convex hull home range
catid = "147"
#subset
panthers_temp <- panthers|> dplyr::filter(CatID == catid)
#initialize
k.int <- round(nrow(sf::st_coordinates(panthers_temp))^0.5,0)
k.search <- seq(k.int, 10*k.int, by=5)
# :'( :'( :'(  :'( :'( :'(  :'( :'( :'(  :'( :'( :'( 
this_works_really_well = FALSE
if(this_works_really_well){
  # :'( :'( :'(  :'( :'( :'(  :'( :'( :'(  :'( :'( :'( 
  #Parameter search for locoh-k
  LoCoH.k.range <- adehabitatHR::LoCoH.k.area(
    sf::as_Spatial(panthers_temp$geometry)
    , unout="km2"
    , krange=k.search
  )
  # local convex hull
  local_convex_hull <- adehabitatHR::LoCoH.k(
    sf::as_Spatial(panthers_temp$geometry)
    , k=k.search[14]
    # , unin = "m"
    # , unout = "km"
    , duplicates = "remove"
  )
}else{
  lch_panther_all <- sf::st_read("../data/lchp_shapefile/lchp_with_holes_all_cats.shp")
  ggplot() + geom_sf(data = lch_panther_all, aes(fill=as.factor(id)))
}
#plot
plt_local_convex_hull <- ggplot() +
  geom_raster(
    data = land_sub_df
    , mapping = aes(x=x,y=y,fill=landcover_class)
  ) +
  geom_sf(data = panthers, mapping = aes(color = CatID), alpha = 0.5, show.legend = F) +
  # plot home range
  geom_sf(
    data = lch_panther_all
    , mapping = aes(color = as.factor(id))
    , fill = "gray"
    , alpha = 0.1
    , lwd = 1
  ) +
  scale_color_viridis_d(option = "magma") +
  scale_fill_viridis_d(option = "turbo", alpha = 0.8) +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0)) +
  labs(
    subtitle = paste0("local convex hull home range")
    , x = "Longitude"
    , y = "Latitude"
    , color = "Cat ID"
    , fill = "Landcover Class"
  ) +
  theme_light() +
  theme(
    legend.position = "none"
    , legend.key.size = unit(0.25, 'cm')
    , legend.text = element_text(size = 6)
    , legend.title = element_text(size = 6)
    , axis.title = element_text(size = 7)
    , axis.text.y = element_text(size=6)
    , axis.text.x = element_text(size=6)
    , panel.border = element_blank()
    , plot.subtitle = element_text(size=9)
    , panel.grid = element_blank()
  ) 
plt_local_convex_hull
```

### brownian bridge home range

```{r, fig.show='hide'}
#Re-format Juldate information:
#function for taking characters of a string from rightmost value
substrRight <- function(x, n){
  substr(x, nchar(x)-n+1, nchar(x))
}

#re-format
panthers <- panthers |> 
  dplyr::mutate(
    Date = substrRight(as.character(Juldate), 3) |> 
      as.numeric() |> 
      as.Date(origin = as.Date("2006-01-01")) |> 
      as.POSIXct(format="%Y-%m-%d", tz = "EST")
  )
#make trajectory object
panther.ltraj <- adehabitatLT::as.ltraj(
  xy=sf::st_coordinates(panthers$geometry)
  , date=panthers$Date
  , id=panthers$CatID
  , typeII=T
)
plot(panther.ltraj)
#telemetry error (average)
sigma2 <- 450
#estimate sigma1
sigma1 <- adehabitatHR::liker(panther.ltraj, sig2 = sigma2, rangesig1 = c(2, 100))
#all panthers
sig1 <- c(
  sigma1[[1]]$sig1
  , sigma1[[2]]$sig1
  , sigma1[[3]]$sig1
  , sigma1[[4]]$sig1
  , sigma1[[5]]$sig1
  , sigma1[[6]]$sig1
)
bb.panther <- adehabitatHR::kernelbb(
  panther.ltraj
  , sig1 = sig1
  , sig2 = sigma2
  , grid = 200
)
# contour
bb.95 <- adehabitatHR::getverticeshr(bb.panther, percent=95) |> 
  sf::st_as_sf() |> 
  sf::st_set_crs(reproject_crs) |>
  dplyr::rename(CatID = id)
# plot
#plot
plt_brownian_bridge <- ggplot() +
  geom_raster(
    data = land_sub_df
    , mapping = aes(x=x,y=y,fill=landcover_class)
  ) +
  geom_sf(data = panthers, mapping = aes(color = CatID), alpha = 0.5, show.legend = F) +
  # plot home range
  geom_sf(
    data = bb.95
    , mapping = aes(color = CatID)
    , fill = "gray"
    , alpha = 0.1
    , lwd = 1
  ) +
  scale_color_viridis_d(option = "magma") +
  scale_fill_viridis_d(option = "turbo", alpha = 0.8) +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0)) +
  labs(
    subtitle = paste0("brownian bridge home range")
    , x = "Longitude"
    , y = "Latitude"
    , color = "Cat ID"
    , fill = "Landcover Class"
  ) +
  theme_light() +
  theme(
    legend.position = "none"
    , legend.key.size = unit(0.25, 'cm')
    , legend.text = element_text(size = 6)
    , legend.title = element_text(size = 6)
    , axis.title = element_text(size = 7)
    , axis.text.y = element_text(size=6)
    , axis.text.x = element_text(size=6)
    , panel.border = element_blank()
    , plot.subtitle = element_text(size=9)
    , panel.grid = element_blank()
  ) 
plt_brownian_bridge
```

### plot grid

minimum convex polygon, fixed bivariate kernel home range, local convex hull home range, and Brownian bridge

```{r}
cowplot::plot_grid(
  plotlist = list(
    plt_minimum_convex_polygon
    , plt_fixed_bivariate_kernel
    , plt_local_convex_hull
    , plt_brownian_bridge
  )
  , ncol = 2
  , rel_heights = c(1,0.8)
)
```


## Question 6

If you were tasked with presenting the home range of these animals to the Florida Fish and Wildlife Commission, which would you select and why? (2 pts)

```{r}

```

# Florida panther resource selection 

## Question 7

Within the Fletcher text, there is a distinction made between Design II and Design III resource selection sampling. What is the difference? (4 pts)

## Question 8

Lines 343 reformats the "use" dataframe. What specially does it do? (2 pts)

`useCatID <- dcast(use, CatID~landcover, length, value.var="CatID")`

## Question 9

Calculate the Design II and Design III resource selection models. (12 pts)

```{r}

```

### part a)
For Design II, are all panthers using the habitats identically? Hint, explore the output of the "widesII".

### part b)
For Design II, which habitats are showing a positive selection? Do any habitats show a significant positive association?

### part c)
For Design III, which habitats are showing a positive selection? Do you see negative associations? 

### part d)
What differences do you see comparing the Design II and Design III global results for habitat selection? 

