---
title: "Lab 7 - Space Use and Resource Selection"
subtitle: "ECOL 620 - Applications in Landscape Ecology"
author: "George Woolsey"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document:
    toc: true
    toc_float: true
    # code_folding: hide
editor_options: 
  chunk_output_type: console
knit: (function(inputFile, encoding){ 
    out_dir <- '../';
    rmarkdown::render(inputFile, encoding = encoding, output_file=file.path(dirname(inputFile), out_dir, 'index.html'));
    file.copy(from = "..//index.html", to = '..///data//lab7_george_woolsey.html', overwrite = TRUE)
  })
---

# Setup

```{r, include=FALSE, warning=F, message=F}
# knit options
knitr::opts_chunk$set(
  echo = TRUE
  , warning = FALSE
  , message = FALSE
  # , results='hide'
  , fig.width = 10
  , fig.height = 7
)
```

```{r}
# bread-and-butter
library(tidyverse)
library(lubridate)
library(viridis)
library(scales)
library(latex2exp)
# visualization
library(kableExtra)
library(cowplot)
library(RColorBrewer)
# spatial analysis
library(terra)
library(adehabitatLT)     #for trajectory information; version 0.3.23 used
library(adehabitatHR)     #Home range estimation; version 0.4.15 used
library(adehabitatHS)     #for selection ratios; version 0.3.13 used
# do i have to???
# library(raster)           #for raster covariate data; version 2.6-7 used
# library(rgdal)            #for reading different types of GIS files; version 1.3-4 used
# set seed
set.seed(11)
```

# Rules

1. Using the provided R Markdown template, answer all questions and show your R code where necessary. Note, some questions are just plain text written answers.

2. Complete your assignment using the R markdown file and submit individual assignments to Canvas. Knit your script and submit an .html file on Canvas. Please use  the following naming convention: lab5_firstname_lastname.html (ex. lab7_kyle_horton.html).  Note, we will not grade labs in any other format.

---

# Florida panther home range

```{r, include=FALSE, eval=FALSE}
...notes...

local convex hull map might be difficult...check for shapefile uploaded to canvas
  is the issue with the holes in the polygons? can use st_make_valid?

resource selection has various levels...focus on type 2 and type 3...trying to figure out where the individuals are, are they selecting preferentially basee on what's available?

type 2 is at broad landscape scale, type 3 narrows down to home range
```


## Load Data

```{r, results='hide', fig.show='hide'}
# landcover source: fwc/fnai
land <- terra::rast("../data/panther_landcover.tif")
reproject_crs <- terra::crs(land)
terra::freq(land)
# Add panther data
panthers <- sf::st_read("../data/panthers.shp") |> 
  sf::st_transform(reproject_crs) |> 
  dplyr::select(-c(X,Y)) |> 
  dplyr::mutate(CatID = as.factor(CatID))
#inspect
str(panthers)
# counts
panthers |> 
  sf::st_drop_geometry() |> 
  dplyr::count(CatID)
panthers |> 
  sf::st_drop_geometry() |> 
  dplyr::count(AgeClass)
panthers |> 
  sf::st_drop_geometry() |> 
  dplyr::count(CatID,AgeClass)
# plot
ggplot() + 
  geom_raster(
    data = land |> as.data.frame(xy=T) |> dplyr::rename(covertyp=3)
    , mapping = aes(x=x,y=y,fill=covertyp)
  ) +
  geom_sf(data = panthers, mapping = aes(color = CatID)) +
  scale_fill_viridis_c(alpha = 0.8) +
  scale_color_viridis_d(option = "magma") +
  theme_light()
```

Reclassify the land cover data based on classification matrix

```{r, results='hide', fig.show='hide'}
# reclass
# load reclassification to resistance surface
landcover_reclass <- read.table("../data/landcover reclass.txt", header = T) |>  
  dplyr::rename_with(tolower)
landcover_reclass |>  dplyr::glimpse()
# create landcover surface raster
rcl_matrix <- landcover_reclass |>  dplyr::select(landcover, changeto) |>  as.matrix()
land_sub <- terra::classify(land, rcl = rcl_matrix)
# investigate
land |>  terra::unique() |>  nrow()
land_sub |>  terra::unique() |>  nrow()
plot(land_sub)

#create forested wetlands layer
wetforest <- land_sub
terra::values(wetforest) <- 0
wetforest[land_sub %in% c(9,11)] <- 1
plot(wetforest)
#create forested uplands layer
dryforest <- land_sub
terra::values(dryforest) <- 0
dryforest[land_sub %in% c(10,12)] <- 1
plot(dryforest)
# create moving window calculation of different forest types
fw <- terra::focalMat(land_sub, d = 5000, type = "circle")
dry.focal <- terra::focal(dryforest, w = fw, fun = "sum", na.rm=T)
wet.focal <- terra::focal(wetforest, w = fw, fun = "sum", na.rm=T)
plot(dry.focal)
plot(wet.focal)
```

```{r, warning=FALSE, message=FALSE, echo=FALSE, include=FALSE}
remove(list = ls()[grep("_temp",ls())])
gc()
```

## Question 1

Within the "panthers" SpatialPointsDataFrame, there are three columns of information. Describe what each represents. (3 pts)

<span style="color: teal;">
The "panthers.shp" data includes radiotelemetry data from Florida panthers (*Puma concolor coryi*) in south Florida with the columns **`r paste0(names(panthers |> sf::st_drop_geometry()), collapse = ", ")`**. Data come from six panthers identified by the column `CatID`, three subadults and three adults denoted in the data column `AgeClass`, that were tracked during 2006 with dates identified by the column `Juldate`.
</span>

## Question 2

Describe what lines 102-104 are achieving. What is "fw" and how is it used in the "focal" function? (4 pts)

```{r, eval=FALSE, include=TRUE}
fw <- raster::focalWeight(land_sub, 5000, 'circle')
dry.focal <- raster::focal(dryforest, w=fw, fun="sum", na.rm=T)
wet.focal <- raster::focal(wetforest, w=fw, fun="sum", na.rm=T)
```

<span style="color: teal;">
The `raster::focalWeight` function calculates a "moving window" weight matrix in which the values add up to one, named `fw`, in a 5 km radius (5,000 m) for use in the `raster::focal` function. The `raster::focal` function calculates the focal ("moving window") values for the neighborhood of focal cells using a matrix of weights and sums the weighted amount of each forest type within the 5 km radius.
</span>

## Question 3

Plot the minimum convex polygon for panther #137. Show the raw locations (points) and the 50%, 75%, and 95% minimum convex polygons. See the example on Canvas for reference.  (5 pts)

```{r}
# take panthers and select the column of the data frame that 
  # includes the animal identification number (CatID) to 
  # calculate Minimum Convex Polygons with the mcp function
# :'( :'( :'(  :'( :'( :'(  :'( :'( :'(  :'( :'( :'( 
# have to convert from sf to sp object to use mcp
sp_panthers_temp <- sf::as_Spatial(panthers)
# function to pass to mcp
my_mcp_fn <- function(pct) {
  adehabitatHR::mcp(
    sp_panthers_temp[,"CatID"]
    , percent = pct
  ) |> 
  sf::st_as_sf()
}
# pcts for minimum convex polys
pct_list <- c(95,75,50)
# create sf object with all pcts in one for plotting
min_convex_polys <- pct_list |> 
  purrr::map(my_mcp_fn) |> 
  dplyr::bind_rows(.id = "column_label") |> 
  dplyr::mutate(
    pct_grp = factor(
      column_label
      , labels = paste0(pct_list,"%")
      , ordered = T
    )
  )
```

Plot for panther #137

```{r}
panther_filter_temp <- "137"
# create df of landcover reclass raster
land_sub_df_temp <- land_sub |> 
  terra::crop(
    panthers |> 
      dplyr::filter(CatID == panther_filter_temp) |> 
      sf::st_bbox() |> 
      sf::st_as_sfc() |> 
      sf::st_buffer(5000) |> 
      terra::vect()
  ) |> 
  as.data.frame(xy=T) |> 
  dplyr::rename(landcover_type=3) |> 
  dplyr::left_join(
    landcover_reclass |> 
      dplyr::count(changeto,description2) 
    , by = c("landcover_type"="changeto")
  ) |> 
  dplyr::mutate(
    landcover_class = as.factor(description2)
  )
#plot
ggplot() +
  geom_raster(
    data = land_sub_df_temp
    , mapping = aes(x=x,y=y,fill=landcover_class)
  ) +
  geom_sf(
    data = min_convex_polys |> dplyr::filter(id == panther_filter_temp)
    , mapping = aes(color = pct_grp)
    , fill = "gray"
    , alpha = 0.1
    , lwd = 2
  ) +
  geom_sf(
    data = panthers |> dplyr::filter(CatID == panther_filter_temp)
    , color = "black"
    , fill = "gray85"
    , shape = 21
  ) +
  scale_fill_viridis_d(option = "turbo", alpha = 0.8) +
  scale_color_manual(
    values = RColorBrewer::brewer.pal(n = length(pct_list)*2, name = "Greys") |> tail(length(pct_list))
  ) +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0)) +
  labs(
    subtitle = paste0("Radiotelemetry locations and minimum convex polygons for panther #", panther_filter_temp)
    , x = "Longitude"
    , y = "Latitude"
    , fill = "Landcover Class"
    , color = "Minimum convex polygons\nHome Range %"
  ) +
  theme_light() +
  theme(
    legend.position = c(0.12, 0.2)
    , legend.key.size = unit(0.25, 'cm')
    , legend.text = element_text(size = 7)
    , legend.title = element_text(size = 7)
    , axis.title = element_text(size = 7)
    , axis.text.y = element_text(size=6)
    , axis.text.x = element_text(size=6)
    , panel.border = element_blank()
    , plot.subtitle = element_text(size=9)
    , panel.grid = element_blank()
  )

```

```{r, warning=FALSE, message=FALSE, echo=FALSE, include=FALSE}
remove(list = ls()[grep("_temp",ls())])
gc()
```

## Question 4

Plot the minimum convex polygon for all panthers overlaid on the reclassified landcover raster. Show the 50% and 95% minimum convex polygons. See the example on Canvas for reference. (6 pts)

```{r}

```

## Question 5

For all panthers, create a four-panel plot showing the home ranges as estimated by: minimum convex polygon, fixed bivariate kernel home range, local convex hull home range, and Brownian bridge. For each approach, plot the 95% contours. Hint, for the local convex hull home range, I used the "LoCoH.k" function with a k of 14. (12 pts). 

```{r}

```

## Question 6

If you were tasked with presenting the home range of these animals to the Florida Fish and Wildlife Commission, which would you select and why? (2 pts)

```{r}

```

# Florida panther resource selection 

## Question 7

Within the Fletcher text, there is a distinction made between Design II and Design III resource selection sampling. What is the difference? (4 pts)

## Question 8

Lines 343 reformats the "use" dataframe. What specially does it do? (2 pts)

`useCatID <- dcast(use, CatID~landcover, length, value.var="CatID")`

## Question 9

Calculate the Design II and Design III resource selection models. (12 pts)

```{r}

```

### part a)
For Design II, are all panthers using the habitats identically? Hint, explore the output of the "widesII".

### part b)
For Design II, which habitats are showing a positive selection? Do any habitats show a significant positive association?

### part c)
For Design III, which habitats are showing a positive selection? Do you see negative associations? 

### part d)
What differences do you see comparing the Design II and Design III global results for habitat selection? 

