---
title: "Lab 7 - Space Use and Resource Selection"
author: "ECOL 620 - Applications in Landscape Ecology"
date: "Spring 2023"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

In this lab, we explore a number of common algorithms for estimating animal home ranges and resource selection. We will use GPS tracking data from Florida panthers to motivate these analyses.

Read “Fletcher_Chapter8.pdf” and open “lab8_practice.Rmd” in R. Complete your assignment using the R markdown file and submit individual assignments to Canvas. Knit your script and submit an .html file on Canvas. Please use  the following naming convention: lab7_firstname_lastname.html (ex. lab7_kyle_horton.html).  Note, we will not grade labs in any other format.



```{r}
library(raster)           #for raster covariate data; version 2.6-7 used
library(reshape2)         #for re-formatting data; version 1.4.3 used
library(rgdal)            #for reading different types of GIS files; version 1.3-4 used
library(adehabitatLT)     #for trajectory information; version 0.3.23 used
library(adehabitatHR)     #Home range estimation; version 0.4.15 used
library(adehabitatHS)     #for selection ratios; version 0.3.13 used
library(tidyverse)
```


```{r}
#set working directory where data were downloaded
setwd("/Users/kylehorton/Desktop/ECOL_620/Course_materials/Week10/Lab7/data_for_lab7")
```


```{r}
###################################################
#8.3.2 Prepping the data
###################################################
#label projection for later use
crs.land <- "+proj=aea +lat_0=24 +lon_0=-84 +lat_1=24 +lat_2=31.5 +x_0=400000 +y_0=0 +ellps=GRS80 +datum=WGS84 +units=m +no_defs"
```


```{r}
#landcover source: fwc/fnai
land <- raster("/Users/kylehorton/Desktop/ECOL_620/Course_materials/Week10/Lab7/data_for_lab7/panther_landcover.tif")
plot(land)
```



```{r}
#Add panther data
panthers = readOGR("/Users/kylehorton/Desktop/ECOL_620/Course_materials/Week10/Lab7/data_for_lab7/panthers.shp")
panthers = spTransform(panthers, crs.land)
#the x and y variables are likely in a different CRS, so I have removed them to avoid confusion
panthers=panthers[,3:5]

#inspect
summary(panthers)
unique(panthers$CatID) #the unique cat IDs
head(panthers)
```

```{r}
#plot
plot(land)
points(panthers, col=panthers$CatID)
```

```{r}
unique(land)
#load reclassification table for reclassifying map
classification <- read.table("/Users/kylehorton/Desktop/ECOL_620/Course_materials/Week10/Lab7/data_for_lab7/landcover reclass.txt", header=TRUE)

#inspect
head(classification)
classification$Description    #original classification
classification$Description2   #re-class

#format for reclassify function;
class <- as.matrix(classification[,c(1,3)])
land_sub <- reclassify(land,rcl=class)

#plot
plot(land_sub)

#create forested wetlands layer
wetforest <- land_sub
values(wetforest) <- 0
wetforest[land_sub==9 | land_sub==11] <- 1

#create forested uplands layer
dryforest <- land_sub
values(dryforest) <- 0
dryforest[land_sub==10 | land_sub==12] <- 1

#5 km moving window to get neighborhood proportion
fw <- raster::focalWeight(land_sub, 5000, 'circle')
dry.focal <- focal(dryforest, w=fw, fun="sum", na.rm=T)
wet.focal <- focal(wetforest, w=fw, fun="sum", na.rm=T)

#merge into a single raster stack
layers <- stack(land_sub, wet.focal, dry.focal)
names(layers) <- c("landcover", "wetforest", "dryforest")

#plot
plot(layers)

```



# Home range analysis

## mcp home range
```{r}
#------------------#
#mcp home range
#------------------#

mcp95 <- mcp(panthers[,"CatID"], percent = 95)
mcp50 <- mcp(panthers[,"CatID"], percent = 50)

#inspect
class(mcp95)
head(mcp95@polygons)

#plot
plot(land_sub)
plot(panthers, add=TRUE, col=panthers$CatID)
plot(mcp95, add=TRUE)
plot(mcp50, add=TRUE, col="orange")

```



## fixed bivariate kernel home range
```{r}
#kernel types:
kernel.href.bivar <- kernelUD(panthers[,"CatID"], h="href", kern="bivnorm")
kernel.href.epa <- kernelUD(panthers[,"CatID"], h="href", kern="epa")

#plot
image(kernel.href.bivar)
image(kernel.href.epa)

#alternative plot for first cat
plot(kernel.href.bivar[[1]])
plot(kernel.href.epa[[1]])
```

```{r}
#UD data
head(kernel.href.bivar[[1]]@data)

#h value for bandwidth
kernel.href.bivar[[2]]@h
kernel.href.bivar[[2]]@h$h
```

```{r}
#least-squares cross validation for h, this will yield a warning of convergence
kernel.lscv.bivar <- kernelUD(panthers[,"CatID"], h="LSCV", kern="bivnorm")

#manually adjust h
kernel.bivar.h1000 <- kernelUD(panthers[,"CatID"], h=1000, kern="bivnorm")

#plot first cat
plot(kernel.bivar.h1000[[1]])
plot(kernel.href.bivar[[1]])

#contour maps of activity for fourth cat
plot(kernel.href.bivar[[4]])
contour.data <- as.image.SpatialGridDataFrame(kernel.href.bivar[[4]])
contour(contour.data, nlevels=5, add=TRUE)
```


## local convex hull home range
```{r}
#subset
panther147 <- panthers[panthers$CatID==147, ]
panther100 <- panthers[panthers$CatID==100, ]

#initialize
k.int <- round(nrow(coordinates(panther147))^0.5,0)
a.int <- round(max(dist(coordinates(panther147))),0)

k.search <- seq(k.int, 10*k.int, by=5)
a.search <- seq(a.int, 2*a.int, by=3000)

#Parameter search for locoh-a
LoCoH.a.range <- LoCoH.a.area(SpatialPoints(coordinates(panther147)), unout="km2", arange=a.search)

#Parameter search for locoh-k
LoCoH.k.range <- LoCoH.k.area(SpatialPoints(coordinates(panther147)), unout="km2", krange=k.search)
```


```{r}
#plot
plot(LoCoH.a.range)
plot(LoCoH.k.range)
```


```{r}
#inspect
a.search[5]
k.search[11]

#re-fit model
LoCoH.k.61 <- LoCoH.k(SpatialPoints(coordinates(panther147)), k=k.search[11])

#plot
plot(LoCoH.k.61)
```


```{r}
#re-fit model
LoCoH.a.100062 <- LoCoH.a(SpatialPoints(coordinates(panther147)), a=a.search[5])
class(LoCoH.a.100062)
#plot
plot(LoCoH.a.100062)
```



## brownian bridge home range
```{r}
#Re-format Juldate information:
#function for taking characters of a string from rightmost value
substrRight <- function(x, n){
  substr(x, nchar(x)-n+1, nchar(x))
}

#re-format
panthers$Juldate <- as.character(panthers$Juldate)
panther.date <- as.numeric(substrRight(panthers$Juldate, 3))
panthers$Date <-as.Date(panther.date, origin=as.Date("2006-01-01"))

#convert to POSIXct object
panthers$Date <- as.POSIXct(panthers$Date,"%Y-%m-%d", tz = "EST")

#convert to data frame
panther.df <- as.data.frame(panthers)

#make trajectory object
panther.ltraj <- as.ltraj(xy=coordinates(panthers), date=panthers$Date, id=panthers$CatID, typeII=T)
```

```{r}
#inspect
head(panther.ltraj)
head(panther.ltraj[[1]], 2)
summary(panther.ltraj)

#plot
plot(panther.ltraj)
```

```{r}
#telemetry error (average)
sigma2 <- 450

#estimate sigma1
sigma1 <- liker(panther.ltraj, sig2 = sigma2, rangesig1 = c(2, 100))

#inspect
#sigma1
```


```{r}
#brownian bridge for Cat 147
bb.147 <- kernelbb(panther.ltraj[6], sig1 = 7.2, sig2 = sigma2, grid = 200)

#all panthers
sig1 <- c(sigma1[[1]]$sig1, sigma1[[2]]$sig1, sigma1[[3]]$sig1, sigma1[[4]]$sig1, sigma1[[5]]$sig1, sigma1[[6]]$sig1)
bb.panther <- kernelbb(panther.ltraj, sig1 = sig1, sig2 = sigma2, grid = 200)
```


```{r}
#plot
plot(panther.ltraj[6])
plot(bb.147)
```


# contrast estimates
```{r}
#home range area estimates
kernel.95 <- getverticeshr(kernel.href.bivar, percent=95)
bb.95 <- getverticeshr(bb.panther, percent=95)

#contrast area
mcp95$area
kernel.95$area
bb.95$area

#plot
par(mfrow=c(1,2))  #sets up the graph window to store two graphs
plot(land_sub)
plot(kernel.95, add=TRUE, col=kernel.95$id)

plot(land_sub)
plot(mcp95, add=TRUE, col=kernel.95$id)

#write to shapefile if needed
#writePolyShape(mcp95, "homerange")
```

########################################################


# Resource Selection

## Point selection functions
```{r}
#use data
use <- raster::extract(layers, panthers)
use <- data.frame(use)

#inspect
head(use)
str(use)

#add CatID
use$CatID <- as.factor(panthers$CatID)

#reformat
useCatID <- dcast(use, CatID~landcover, length, value.var="CatID")

#inspect
useCatID

#add land-cover names
newclass.names <- unique(classification[,3:4])
names(useCatID) <- c("CatID", as.character(newclass.names[1:13,2]))

#inspect
useCatID
```


## design II availability: population availability
```{r}
#get availability points
set.seed(8)
rand.II <- sampleRandom(layers, size=1000)
rand.II <- data.frame(rand.II)

#inspect
head(rand.II)
str(rand.II)

rand.II.land <- as.factor(rand.II$landcover)

#get counts of each landcover type
avail.II <- tapply(rand.II.land, rand.II.land, length)

#inspect
avail.II
```


```{r}
#add land-cover names
names(avail.II) <- as.character(newclass.names[1:14,2])

#inspect
avail.II

#remove exotics, which were not observed in use sample
avail.II <- avail.II[c(-14)]
```

## design III availability: within home-range availability for each individual
```{r}
cat.unique <- unique(panthers$CatID)
samples <- 200
rand.III <- matrix(nrow=0, ncol=4)

#loop for all individuals
for(i in 1:length(cat.unique)){

  id.i <- cat.unique[i]
  cat.i <- panthers[panthers$CatID==id.i,]
  mcp.i <- mcp(SpatialPoints(coordinates(cat.i)), percent = 99)
  rand.i <- spsample(mcp.i, type="random", n=samples)
  rand.i.sample <- raster::extract(layers, rand.i)

  #make a matrix of CatID and rand samples
  cat.i <- as.numeric(rep(cat.unique[i], length(rand.i)))
  rand.cat.i <- cbind(cat.i, rand.i.sample)
  rand.III <- rbind(rand.III, rand.cat.i)
}
```


```{r}
#inspect
head(rand.III)
class(rand.III)
str(rand.III)
```


```{r}
#reshape data
rand.III <- data.frame(rand.III)
rand.III$cat.i <- as.factor(rand.III$cat.i)
avail.III <- dcast(rand.III, cat.i~landcover, length, value.var="cat.i")

names(avail.III)[2:14] <- as.character(newclass.names[1:13,2])
#inspect
avail.III
```

## selection ratios

### Design II:
```{r}
sel.ratioII <- widesII(u=useCatID[,c(2:ncol(useCatID))], a=as.vector(avail.II), avknown=FALSE, alpha = 0.05)

#inspect
sel.ratioII
sel.ratioII$wi
sel.ratioII$se.wi
```


```{r}
#plot
#opar <- par(mfrow = c(2,2))
#plot(sel.ratioII, errbar = c("CI"))
#par(opar)

ggplot()+
  geom_point(aes(y=sel.ratioII$wi, x=row.names(as.data.frame(sel.ratioII$wi))))+
  theme_classic()+
  geom_hline(yintercept = 1)+
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
  geom_errorbar(aes(ymin=sel.ratioII$ICwiupper,ymax=sel.ratioII$ICwilower, x=row.names(as.data.frame(sel.ratioII$wi)),width=0.2))+
  labs(x="Covertype", y=  "Manly selectivity measure\n (used/available)")
```


### Design III:
```{r}
sel.ratioIII <- widesIII(u=useCatID[,c(2:ncol(useCatID))], a=avail.III[,2:14], avknown=FALSE, alpha = 0.05)

#inspect
sel.ratioIII
sel.ratioIII$wi
sel.ratioIII$se.wi
sel.ratioIII$ICwiupper
sel.ratioIII$ICwilower
```


```{r}
ggplot()+
  geom_point(aes(y=sel.ratioIII$wi, x=row.names(as.data.frame(sel.ratioIII$wi))))+
  theme_classic()+
  geom_hline(yintercept = 1)+
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
  geom_errorbar(aes(ymin=sel.ratioIII$ICwiupper,ymax=sel.ratioIII$ICwilower, x=row.names(as.data.frame(sel.ratioIII$wi)),width=0.2))+
  labs(x="Covertype", y=  "Manly selectivity measure\n (used/available)")
```

